---
title: "Test CBE State vs GHGI"
output: github_document
---

```{r setup, include=FALSE}
source("../R/utils.R")
source("../R/StateEEIOFigures.R")
source("../R/StateEEIOCalculations.R")
library(useeior)
library(knitr)
soi <- "CA"
```

To investigate issue [#1](https://github.com/cornerstone-data/USEEIO-State/issues/1), this analysis compares the
California CBE as presented and compares it to the CA GHGI 2019 used in the model.

Download and read in the CA model for 2019 with GHGs.
```{r getCAmodel, echo=TRUE}
#rmarkdown::render('../examples/DownloadandSaveModelLocally.Rmd', params =  list(states = c('CA'),model="v1.1-GHG",years=c(2019)))
model <- readRDS("../models/CAEEIOv1.1-GHG-19.rds")
```

## Validate the model
Make sure the model passes existing validation tests.
```{r validateModel}
printValidationResults(model)
```


## Compare CBE from CA as done in report vs GHGI
First calculate the CBE as done in the report, using the final perspective which is the default
for calculateStateCBE. 

```{r calcCBEdefault, echo=TRUE}
cbe <- calculateStateCBE(model,perspective="FINAL")
cbe_fromCA <- subset(cbe, endsWith(rownames(cbe), soi)) 
```

Total California CBE in 2019 from CA calculated with the final perspective
is `r sum(cbe_fromCA)` metric tons CO2e. 
The total from all locations is `r sum(cbe)` metric tons CO2e.

```{r getGHGI}
ghgi <- getStateGHGI(model)
total_ghgi <- sum(ghgi[,"Greenhouse Gases"])
```

The total California GHGI from model GHGI data for 2019 is `r total_ghgi` metric tons CO2e.

The result is that the CBE attributed to CA is greater than the GHGI for CA.

## Change in Method - Use DIRECT Perspective

The DIRECT perspective calculation actually reveals emissions where they occur,
rather than attributing them to the final goods and services consumed in the state.
Recalculate the CBE using the direct perspective.

```{r calculateCBEdirect, echo=TRUE}

cbe <- calculateStateCBE(model,perspective="DIRECT")
cbe_fromCA <- subset(cbe, endsWith(rownames(cbe), "CA")) 

```

Total California CBE from CA in 2019 calculated with the direct perspective is `r sum(cbe_fromCA)` metric tons CO2e.
The total calculated from all locations is `r sum(cbe)` metric tons CO2e.

To compare the CBE to the GHGI, the GHGI has to be translated into commodity format
We use the commmodity mix matrix to do this.

```{r compareCBEGHGIdirect, echo=TRUE}

#Get the model list of industries in CA
CA_inds <- model$Industries$Code_Loc[endsWith(model$Industries$Code_Loc, "CA")]

#Index GHGI by CA industries
mghgi <- ghgi[match(CA_inds, rownames(ghgi)), , drop=FALSE]


print("Are the industry rows in the same order in GHGI and in CBE?")
identical(rownames(mghgi),CA_inds)
#mghgi<- as.matrix(ghgi)
#dim(mghgi)
C_m_CA <- subset(model$C_m, endsWith(rownames(model$C_m), "CA") , endsWith(colnames(model$C_m), "CA"))
#dim(C_m_CA)

#Left multiply C_m_CA by mghgi to get GHGI in commodity format
ghgi_c <- C_m_CA %*% mghgi

#Add household back in
ghgi_c <- rbind(ghgi_c,"F010/US-CA"=ghgi["F010/US-CA","Greenhouse Gases"])

#Create a table to compare CBE and GHGI by sector
compare <- merge(cbe_fromCA, ghgi_c, by=0)
colnames(compare) <- c("Sector", "CBE_fromCA", "GHGI")
compare$CBE_frac_of_GHGI <- compare$'CBE_fromCA'/compare$GHGI
rownames(compare) <- rownames(cbe_fromCA)

kable(compare, caption = "Comparison of CBE from CA using direct perspective vs GHGI")

```

Show the commodit
```{r comswithhigherGHGs, echo=FALSE}
higher_ghg <- subset(compare, CBE_frac_of_GHGI > 1,c("CBE_frac_of_GHGI"), drop=FALSE)
kable(higher_ghg, caption = "Commodities with CBE greater than GHGI")
```



## Try it with domestic requirements only

```{r calculateCBEdirectwithdomestic, echo=FALSE}

cbe <- calculateStateCBE(model,perspective="DIRECT", domestic=TRUE)

cbe_fromCA <- subset(cbe, endsWith(rownames(cbe), "CA")) 

```

Total California CBE from CA in 2019 calculated with the direct perspective with domestic requirements is `r sum(cbe_fromCA)` metric tons CO2e.
The total calculated from all locations is `r sum(cbe)` metric tons CO2e.
This is the same as without domestic requirements.

```{r compareCBEGHGIdirectonlydomestic, echo=FALSE}

compare <- merge(cbe_fromCA, ghgi_c, by=0, how="right")

colnames(compare) <- c("Sector", "CBE_fromCA", "GHGI")

compare$CBE_frac_of_GHGI <- compare$'CBE_fromCA'/compare$GHGI

kable(compare, caption = "Comparison of CBE from CA using direct perspective and domestic inputs only vs GHGI")


```

To take a clos
We can also compare the total uses of CA commodities with the total commodity output (q). The total uses should not be greater than q.

```{r compareUsesDirect, echo=FALSE}
#
demand_vectors <- model$DemandVectors$vectors
prod <- data.frame(demand_vectors$`2019_US-CA_Production_Domestic`)
cons <- data.frame(demand_vectors$`2019_US-CA_Consumption_Domestic`)


prod_soi <- subset(prod, endsWith(rownames(prod), soi))
cons_soi <- subset(cons, endsWith(rownames(cons), soi))

q <- matrix(model$q,dimnames = c(list(names(model$q), "q")))
q_soi <-  subset(q, endsWith(rownames(q), soi), drop=FALSE)

intermediate_uses <- getStateUsebyType(model, type="intermediate", domestic=TRUE)
intermediate_uses_soi <- subset(intermediate_uses, endsWith(rownames(intermediate_uses), soi))

compare_use <- cbind(prod_soi, cons_soi, intermediate_uses_soi, q_soi)
colnames(compare_use) <- c("Production_Demand", "Consumption_Demand", "Intermediate_Uses", "q")

compare_use$q_calc_consumption <- rowSums(compare_use[,c("Intermediate_Uses","Consumption_Demand")])
compare_use$cons_frac_of_q <- compare_use$q_calc_consumption/compare_use$q 

compare_use$q_calc_production <- rowSums(compare_use[,c("Intermediate_Uses","Production_Demand")])
compare_use$prod_frac_of_q <- compare_use$q_calc_production/compare_use$q


kable(compare_use, caption = "Comparison of Uses to Commodity output (q) for production and consumption with domestic requirements only")
```

Some of the commodities have uses greater than q. Show these.
```{r higherUses, echo=FALSE}

higher_uses <- subset(compare_use, cons_frac_of_q > 1,c("cons_frac_of_q"), drop=FALSE)
kable(higher_uses, caption = "Commodities with Uses greater than output (q)")


```

These require further investigation.


